### 2015-10-16 sample-paper/sample-paper.tex
a simple trix exampling referencing in latex, with latex command direct in pandoc markdown

\documentclass[a4paper,12pt]{article}
\usepackage{geometry}
\geometry{left=3cm,right=3cm,top=3cm,bottom=3cm,footskip=2cm}
\usepackage{setspace}	
\linespread{1.3}

\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Donald Duck},
            pdftitle={A sample paper},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\title{A sample paper}
\author{Donald Duck}
\date{October 1, 2014}

\begin{document}
\maketitle
\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
tempor incididunt ut labore et dolore magna aliqua. Ut enimad minim
veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
occaecat cupidatat non proident, sunt in culpa qui officia deserunt
mollit anim id est laborum.
\end{abstract}

\section{Section 1}\label{section-1}

\subsection{Subsection 1.1}\label{subsection-1.1}

Lorem \emph{ipsum} \textbf{dolor} sit amet, consectetur adipisicing
elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi
ut aliquip ex ea commodo consequat.

\subsection{Subsection 1.2}\label{subsection-1.2}

Sed ut perspiciatis unde omnis iste natus error sit voluptatem
accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab
illo inventore veritatis et quasi architecto beatae vitae dicta sunt
explicabo.

\begin{itemize}
%\tightlist
\item
  one item
\item
  two items
\item
  three items
\item
  four items
\end{itemize}

Sed ut perspiciatis unde omnis iste natus error sit voluptatem
accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab
illo inventore veritatis et quasi architecto beatae vitae dicta sunt
explicabo.

You can enter LaTeX equations as inline math: such as
\(({e}^{i\pi }+1=0)\) or:

\(\mathbf{V}_1 \times \mathbf{V}_2 = \begin{vmatrix} \mathbf{i} & \mathbf{j} & \mathbf{k} \\ \frac{\partial X}{\partial u} & \frac{\partial Y}{\partial u} & 0 \\ \frac{\partial X}{\partial v} & \frac{\partial Y}{\partial v} & 0 \end{vmatrix}\)

\section{Section 2}\label{section-2}

\subsection{Subsection 2.1}\label{subsection-2.1}

Sed ut perspiciatis unde omnis iste natus error sit voluptatem
accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab
illo inventore veritatis et quasi architecto beatae vitae dicta sunt
explicabo.

Sed ut perspiciatis unde omnis iste natus error sit voluptatem
accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab
illo inventore veritatis et quasi architecto beatae vitae dicta sunt
explicabo \href{http://www.http://daringfireball.net}{link}. You enter
citations using @ and bibtex citation key, such as Duck (2014) if you
want only the year in brackets, or (Duck 2013) for a normal ciation in
author-date form within brackets.

This is the code for an image. Much simpler than in LaTeX. But you can
also use LaTeX code here. Just keep in mind that LaTeX code is only
rendered when converted to PDF. Conversion to Word or html won't pick it
up. And see that figure referenced see \ref{fig:mylabel}.

\begin{figure}[htbp]
\centering
\includegraphics{sample-image.jpg}
\caption{image \label{fig:mylabel} caption}
\end{figure}

\section{References}\label{references}

The reference list is added here automatically by Pandoc:

\hyperdef{}{ref-donaldduck2013}{\label{ref-donaldduck2013}}
Duck, Donald. 2013. ``Turning Nothing into Money.'' \emph{Money
Transactions} 15: 205--246.

\hyperdef{}{ref-donaldduck2014}{\label{ref-donaldduck2014}}
---------. 2014. ``Turning Something Else into Money.'' \emph{Money
Money Journal} 5: 101--141.

\end{document}